Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary

### 2020-05-19 scripts/MouseSexDiffApril20_2020.Rmd
More supplements stuff. Added in mean-variance relationships

---
title: "Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications"
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Alistair
  Senior, Daniel Noble, & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  bookdown::html_document2: 
            toc: true
            code_folding: hide
            toc_depth: 6
            toc_float: yes
            number_sections: yes
            theme: flatly
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

```{r packages1, echo = FALSE, eval = TRUE, messages = FALSE}
library(pacman)
pacman::p_load(readr, dplyr,metafor, devtools, purrr, tidyverse, tidyr, tibble, kableExtra, robumeta, ggpubr, ggplot2, png, grid, here, knitr)
```

@@

This supplementary material follows the general structure described in our methods and is outlined here in Figure \@ref(fig:FigG1). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis. This iss followed by how the first order and second order meta-analsyes were conducted. Supplemental tables and figures that are referenecd within the main manuscript can be found at the end of the document. Various parts of the workflow can be conviently referenecd using the side table of contents panel. We will refer back to this figure as we describe the various aspects of the code and data processing in relevant places.

```{r FigG1, fig.cap="Workflow of data process and meta-analysis"}

@@

calculate_population_stats <- function(mydata, min_individuals = 5){
    mydata %>% 
    group_by(population,
             production_center,  
             strain_name,  
             sex) %>%
    summarise(trait  = parameter_name[1],
              x_bar  = mean(data_point, na.rm = TRUE),
              x_sd   =   sd(data_point, na.rm = TRUE),
              n_ind  = n()) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%   # NOTE here that you are excluding data with 5 exactly....
    group_by(population) %>%
    filter(all(c("male", "female") %in% sex)) %>% # Removes any entries where only 1 sex is present
    multi_spread(sex, c(x_bar, x_sd, n_ind))      # Spreads the data out into wide format, so sexes are beside each other which is more typical of meta-analytic data
}

@@

@@

@@

In this section, we cover the workflow described in Figure \@ref(fig:FigG1) C and D.

@@

This section covers Figure \@ref(fig:FigG1) C, by conducting a meta-analsyis for all traits and producing the meta-analysed effect sizes (lnVR, lnCVR and lnRR).

@@

Here we cover Figure \@ref(fig:FigG1) D, removing non-coverging model results and non-informative traits.Procedure names, grouping variables and trait names ("parameter_names") are the merged back together with the results from the metafor analysis above.

@@

## Meta-analysis: condensing non-independent traits 
This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted meta analyses on these correlated parameters to collapse the number of levels. This section describes the workflow depicted in Figure \@ref(fig:FigG1) E & F.

### Collapsing and merging correlated parameters, preparation for Table S2
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation. The workflow here relates to Figure \@ref(fig:FigG1) E.

@@

## Second-order meta-analysis for functional groups
Here, we  describe the workflow depicted in Figure \@ref(fig:FigG1) H. We now use our uncorrelated traits from our first order meta-analysis to conduct a secondary meta-analysis by analysing the 9 functional trait groups. We esatimte an overall mean lnCVR, lnVR and lnRR for each functional trait group.

@@

The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses and specifically Figure \@ref(fig:FigG1) I & J. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

@@

Here we exclude traits without variation between mouse strains and merge data sets containing metafor results with procedure etc. names. Dealing with the correlated parameters, and running the second -order meta-analysis for heterogeneity data (Figure \@ref(fig:FigG1) I)

@@

# Meta-analysis results
## Preparation for Figure 4 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

## Figure 4

Panel A shows the numbers of traits across functional groups that are either male-biased (blue-green) or female-biased (orange-red), as calculated in Figure \@ref(fig:FigG1) D. Panel B shows effect sizes and 95% CI from separate meta-analysis for each functional group (step H in Figure \@ref(fig:FigG1)). Both panels represent results evaluated across all traits (Phase 3, Figure \@ref(fig:FigG1)). Traits that are male biased are shown in blue, whereas female bias data is represented in orange.

```{r Fig4}
Fig4 <- ggarrange(malebias_Fig2_alltraits, Metameta_Fig3_alltraits,  nrow = 2, align = "v", heights = c(10, 9), labels = c("A", "B"))
Fig4
```

## Preparation for Figure 5
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

To do this, we select only traits that have CIs that do not overlap with zero.
The code below creates Figure 5A.
```{r Fig5_prep}

 meta.plot2.sig <- meta_clean %>%
                   mutate(lnCVRsig = ifelse(lnCVR_lower * lnCVR_upper > 0, 1, 0), 
                           lnVRsig = ifelse( lnVR_lower * lnVR_upper  > 0, 1, 0),
                           lnRRsig = ifelse( lnRR_lower * lnRR_upper  > 0, 1, 0))

    meta.plot2.sig.c <- gather(meta.plot2.sig.b, trait, value, lnCVR:lnRR)
meta.plot2.sig.c$sig <- "placeholder"

  meta.plot2.sig.c$sig <- ifelse(meta.plot2.sig.c$trait == "lnCVR", 
                                 meta.plot2.sig.c$lnCVRsig,
                                 ifelse(meta.plot2.sig.c$trait == "lnVR", 
                                        meta.plot2.sig.c$lnVRsig, 
                                        meta.plot2.sig.c$lnRRsig))

# Choosing sex biased ln-ratios significantly larger than 0
meta.plot2.sig.malebias <- meta.plot2.sig.c %>%
                           group_by_at(vars(trait, GroupingTerm)) %>%
                           filter(sig == 1) %>%
                           summarise(male_sig = sum(value > 0), female_sig = sum(value < 0), total = male_sig + female_sig)

meta.plot2.sig.malebias <- ungroup(meta.plot2.sig.malebias) %>%
                          # add "Hearing" for lnCVR (not filtered as only zeros)
                          add_row(trait = "lnCVR", GroupingTerm = "Hearing", male_sig = 0, female_sig = 0, .before = 4) %>% 
                          mutate(malepercent = male_sig * 100 / total, femalepercent = female_sig * 100 / total)

  meta.plot2.sig.bothsexes <- as.data.frame(meta.plot2.sig.malebias)
meta.plot2.sig.bothsexes.b <- gather(meta.plot2.sig.bothsexes, key = sex, value = percent, malepercent:femalepercent, factor_key = TRUE)

meta.plot2.sig.bothsexes.c <- meta.plot2.sig.bothsexes.b %>% 
                              group_by(trait, sex) %>% 
                              summarise(GroupingTerm = "All", 
                                        male_sig = sum(male_sig), 
                                        female_sig = sum(female_sig), 
                                        total = male_sig + female_sig, 
                                        label = "CI not overlapping zero", 
                                        samplesize = sum(samplesize)) %>%
                              mutate(percent = ifelse(sex == "femalepercent", 
                                                      female_sig*100/(male_sig+female_sig), 
                                                      male_sig*100/(male_sig+female_sig))) %>%
                              bind_rows(meta.plot2.sig.bothsexes.b, .) %>%
                              mutate(rownumber = row_number()) %>%
                              .[c(37, 1:9, 39, 10:18, 38, 19:27, 40, 28:36), ]

  #line references in previous code line corresponding to: 
  #'lnCVR(male(All)), lnCVR(male('single grouping terms'), lnRR(male(All)), lnRR(male('single grouping terms')),
  #lnCVR(female(All)), lnCVR(female('single grouping terms'), lnRR(female(All)), lnRR(female('single grouping terms'))'

meta.plot2.sig.bothsexes.c$GroupingTerm <- factor(meta.plot2.sig.bothsexes.c$GroupingTerm,
                                                  levels = c("Behaviour", "Morphology", "Metabolism", "Physiology", 
                                                             "Immunology", "Hematology", "Heart", "Hearing", "Eye", "All")) 
meta.plot2.sig.bothsexes.c$GroupingTerm <- factor(meta.plot2.sig.bothsexes.c$GroupingTerm, 
                                                  rev(levels(meta.plot2.sig.bothsexes.c$GroupingTerm)))

# Plot Fig2 all significant results (CI not overlapping zero)
# Several grouping terms are added post-hoc (with no data to display): no significant lnCVR for 'Hearing' in either sex; no sig. male-biased lnCVR for 'Immunology' and 'Eye, and no significant female-biased lnRR for 'Eye'.

malebias_Fig2_sigtraits <-
  ggplot(meta.plot2.sig.bothsexes.c) +
  aes(x = GroupingTerm, y = percent, fill = sex) +
  ylab("Percentage") +
  geom_col() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "gray40") +
  geom_text(
    data = subset(meta.plot2.sig.bothsexes.c, samplesize != 0), aes(label = samplesize), position = position_stack(vjust = .5),
    color = "white", size = 3.5) +
  facet_grid(cols = vars(trait), 
             rows = vars(label), 
             labeller = label_wrap_gen(width = 18),
             scales = "free", space = "free") +
  scale_fill_brewer(palette = "Set2") +
  theme_bw(base_size = 18) +
  theme(strip.text.y = element_text(angle = 270, size = 10, margin = margin(t = 15, r = 15, b = 15, l = 15)),
        strip.text.x = element_text(size = 12),
        strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
        text = element_text(size = 14),
        panel.spacing = unit(0.5, "lines"),
        panel.border = element_blank(),
        axis.line = element_line(),
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.title.y = element_blank()) +
  coord_flip()

#### Figure 5 B - traits with CI not overlapping 0 
# Prepare data 
# create column with 1 = different from zero, 0 = zero included in CI

meta.male.plot3.sig <- metacombo %>%
                       mutate(sigCVR = ifelse(lnCVR_lower > 0, 1, 0),
                               sigVR = ifelse( lnVR_lower > 0, 1, 0),
                               sigRR = ifelse( lnRR_lower > 0, 1, 0))

# Significant subset for lnCVR
metacombo_male.plot3.CVR <- meta.male.plot3.sig    %>%
                            filter(sigCVR == 1)    %>%
                            group_by(GroupingTerm) %>%
                            nest()

metacombo_male.plot3.CVR.all <- meta.male.plot3.sig %>%
                                filter(sigCVR == 1) %>%
                                nest(data = everything())

# Significant subset for lnVR
metacombo_male.plot3.VR <- meta.male.plot3.sig    %>%
                           filter(sigVR == 1)     %>%
                           group_by(GroupingTerm) %>%
                           nest()

metacombo_male.plot3.VR.all <- meta.male.plot3.sig    %>%
                               filter(sigVR == 1)     %>%
                               nest(data = everything())

# Significant subset for lnRR
metacombo_male.plot3.RR <- meta.male.plot3.sig    %>%
                           filter(sigRR == 1)     %>%
                           group_by(GroupingTerm) %>%
                           nest()

metacombo_male.plot3.RR.all <- meta.male.plot3.sig %>%
                               filter(sigRR == 1)  %>%
                               nest(data = everything())

plot3.male.meta.CVR <- metacombo_male.plot3.CVR %>%
                       mutate(model_lnCVR = map(data, 
                                                ~ metafor::rma.uni(yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96), 
                                              control = list(optimizer = "optim", 
                                                             optmethod = "Nelder-Mead", 
                                                             maxit = 1000), 
                                              verbose = F)))

plot3.male.meta.VR <- metacombo_male.plot3.VR %>%
                      mutate(model_lnVR = map(data, 
                                              ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                                            control = list(optimizer = "optim", 
                                                           optmethod = "Nelder-Mead", 
                                                           maxit = 1000), 
                                            verbose = F)))

plot3.male.meta.RR <- metacombo_male.plot3.RR %>%
                      mutate(model_lnRR = map(data, 
                                              ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                                              control = list(optimizer = "optim", 
                                                             optmethod = "Nelder-Mead", 
                                                             maxit = 1000), 
                                              verbose = F)))

plot3.male.meta.CVR.all <- metacombo_male.plot3.CVR.all %>%
                            mutate(model_lnCVR = map(data,
                                                     ~ metafor::rma.uni(yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
                                                  control = list(optimizer = "optim", 
                                                                 optmethod = "Nelder-Mead", 
                                                                 maxit = 1000), 
                                                  verbose = F)))

plot3.male.meta.CVR.all <- plot3.male.meta.CVR.all %>% 
                            mutate(GroupingTerm = "All")

plot3.male.meta.VR.all <- metacombo_male.plot3.VR.all %>%
                          mutate(model_lnVR = map(data, 
                                                  ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                                                  control = list(optimizer = "optim", 
                                                                 optmethod = "Nelder-Mead", 
                                                                 maxit = 1000), 
                                                  verbose = F)))

plot3.male.meta.VR.all <- plot3.male.meta.VR.all %>% 
                          mutate(GroupingTerm = "All")

plot3.male.meta.RR.all <- metacombo_male.plot3.RR.all %>%
                          mutate(model_lnRR = map(data, 
                                                  ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                                                control = list(optimizer = "optim", 
                                                               optmethod = "Nelder-Mead", 
                                                               maxit = 1000), 
                                                verbose = F)))

plot3.male.meta.RR.all <- plot3.male.meta.RR.all %>% 
                          mutate(GroupingTerm = "All")

plot3.male.meta.CVR <- bind_rows(plot3.male.meta.CVR, plot3.male.meta.CVR.all)
 plot3.male.meta.VR <- bind_rows(plot3.male.meta.VR,   plot3.male.meta.VR.all)
 plot3.male.meta.RR <- bind_rows(plot3.male.meta.RR,   plot3.male.meta.RR.all)

plot3.male.meta.CVR.b <- as.data.frame(plot3.male.meta.CVR %>% 
                                       group_by(GroupingTerm) %>%
                                       mutate( lnCVR = map_dbl(model_lnCVR, pluck(2)), 
                                               lnCVR_lower = map_dbl(model_lnCVR, pluck(6)), 
                                               lnCVR_upper = map_dbl(model_lnCVR, pluck(7)), 
                                               lnCVR_se = map_dbl(model_lnCVR, pluck(3)), 
                                               lnCVR_I2 = map_dbl(model_lnCVR, pluck("I2"))))[, c(1, 4:8)]

add.row.hearing <- as.data.frame(t(c("Hearing", NA, NA, NA, NA, NA))) %>% 
                                  setNames(names(plot3.male.meta.CVR.b))

plot3.male.meta.CVR.b <- bind_rows(plot3.male.meta.CVR.b, add.row.hearing)
plot3.male.meta.CVR.b <- plot3.male.meta.CVR.b[order(plot3.male.meta.CVR.b$GroupingTerm), ]

plot3.male.meta.VR.b <- as.data.frame(plot3.male.meta.VR %>% 
                                      group_by(GroupingTerm) %>%
                                      mutate(lnVR = map_dbl(model_lnVR, pluck(2)), 
                                             lnVR_lower = map_dbl(model_lnVR, pluck(6)),
                                             lnVR_upper = map_dbl(model_lnVR, pluck(7)), 
                                             lnVR_se = map_dbl(model_lnVR, pluck(3)),
                                             lnVR_I2 = map_dbl(model_lnVR, pluck("I2"))))[, c(1, 4:8)]

plot3.male.meta.RR.b <- as.data.frame(plot3.male.meta.RR %>% 
                                      group_by(GroupingTerm) %>% 
                                      mutate(lnRR = map_dbl(model_lnRR, pluck(2)), 
                                             lnRR_lower = map_dbl(model_lnRR, pluck(6)),
                                             lnRR_upper = map_dbl(model_lnRR, pluck(7)), 
                                             lnRR_se = map_dbl(model_lnRR, pluck(3)),
                                             lnRR_I2 = map_dbl(model_lnRR, pluck("I2"))))[, c(1, 4:8)]

overall.male.plot3 <- full_join(plot3.male.meta.CVR.b, plot3.male.meta.VR.b)
overall.male.plot3 <- full_join(overall.male.plot3,    plot3.male.meta.RR.b)

overall.male.plot3$GroupingTerm <- factor(overall.male.plot3$GroupingTerm, 
                                          levels = c("Behaviour", "Morphology", "Metabolism", "Physiology", 
                                                     "Immunology", "Hematology", "Heart", "Hearing", "Eye", "All"))
overall.male.plot3$GroupingTerm <- factor(overall.male.plot3$GroupingTerm, 
                                          rev(levels(overall.male.plot3$GroupingTerm)))

overall.male.plot3$GroupingTerm <- factor(overall.male.plot3$GroupingTerm, 
                                          levels = c("Behaviour", "Morphology", "Metabolism", "Physiology", 
                                                     "Immunology", "Hematology", "Heart", "Hearing", "Eye", "All"))
overall.male.plot3$GroupingTerm <- factor(overall.male.plot3$GroupingTerm, 
                                          rev(levels(overall.male.plot3$GroupingTerm)))

lnCVR.ci <- overall3.male.sig            %>%
            filter(parameter == "lnCVR") %>%
            mutate(ci.low = lnCVR_lower, 
                   ci.high = lnCVR_upper)

lnRR.ci <- overall3.male.sig           %>%
           filter(parameter == "lnRR") %>%
           mutate(ci.low = lnRR_lower, 
                 ci.high = lnRR_upper)

overall4.male.sig <- bind_rows(lnCVR.ci, lnRR.ci) %>% 
                     select(GroupingTerm, 
                            parameter, 
                            value, 
                            ci.low, 
                            ci.high)

Metameta_Fig3_male.sig <- overall4.male.sig %>%
  ggplot(aes(y = GroupingTerm, x = value)) +
  geom_errorbarh(aes(xmin = ci.low,
                     xmax = ci.high),
                 height = 0.1, 
                 show.legend = FALSE) +
  geom_point(aes(shape = parameter),
             fill = "mediumaquamarine", 
             color = "mediumaquamarine", 
             size = 2.2, 
             show.legend = FALSE) +
  scale_x_continuous(limits = c(0, 0.4),
                     breaks = c(0, 0.3),
                       name = "Effect size") +
  geom_vline(xintercept = 0,
                  color = "black",
              linetype = "dashed") +
  facet_grid(cols = vars(parameter), 
             rows = vars(label),
             labeller = label_wrap_gen(width = 23),
             scales = "free",
             space = "free") +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 270, 
                                    size = 10, 
                                    margin = margin(t = 15, r = 15, b = 15, l = 15)),
        strip.text.x = element_text(size = 12),
        strip.background = element_rect(colour = NULL, 
                                        linetype = "blank", 
                                        fill = "gray90"),
        text = element_text(size = 14),
        panel.spacing = unit(0.5, "lines"),
        panel.border = element_blank(),
        axis.line = element_line(),
        panel.grid.major.x = element_line(linetype = "solid", 
                                          colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", 
                                          color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

# as above: Prepare data for traits with CI not overlapping 0
# create column with 1 = different from zero, 0 = zero included in CI

meta.female.plot3.sig <- metacombo %>%
  mutate(sigCVR = ifelse(lnCVR_upper < 0, 1, 0),
          sigVR = ifelse(lnVR_upper < 0, 1, 0),
          sigRR = ifelse(lnRR_upper < 0, 1, 0))

metacombo_female.plot3.CVR <- meta.female.plot3.sig  %>%
                              filter(sigCVR == 1)    %>%
                              group_by(GroupingTerm) %>%
                              nest()

metacombo_female.plot3.CVR.all <- meta.female.plot3.sig %>%
                                  filter(sigCVR == 1)   %>%
                                  nest(data = everything())

metacombo_female.plot3.VR <- meta.female.plot3.sig  %>%
                             filter(sigVR == 1)     %>%
                             group_by(GroupingTerm) %>%
                             nest()

metacombo_female.plot3.VR.all <- meta.female.plot3.sig  %>%
                                 filter(sigVR == 1)     %>%
                                 nest(data = everything())

metacombo_female.plot3.RR <- meta.female.plot3.sig  %>%
                             filter(sigRR == 1)     %>%
                             group_by(GroupingTerm) %>%
                             nest()

metacombo_female.plot3.RR.all <- meta.female.plot3.sig %>%
                                 filter(sigRR == 1)    %>%
                                 nest(data = everything())

plot3.female.meta.CVR <- metacombo_female.plot3.CVR %>%
                         mutate(model_lnCVR = map(data, 
                                                  ~ metafor::rma.uni(yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
                                                  control = list(optimizer = "optim", 
                                                                 optmethod = "Nelder-Mead", 
                                                                 maxit = 1000), 
                                                  verbose = F)))

plot3.female.meta.VR <- metacombo_female.plot3.VR %>%
                        mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                                                control = list(optimizer = "optim", 
                                                               optmethod = "Nelder-Mead", 
                                                               maxit = 1000), 
                                                verbose = F)))

plot3.female.meta.RR <- metacombo_female.plot3.RR %>%
                        mutate(model_lnRR = map(data, ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                                                control = list(optimizer = "optim", 
                                                               optmethod = "Nelder-Mead", 
                                                               maxit = 1000), 
                                                verbose = F)))

plot3.female.meta.CVR.all <- metacombo_female.plot3.CVR.all %>%
                             mutate(model_lnCVR = map(data, ~ metafor::rma.uni(yi = .x$lnCVR, sei=(.x$lnCVR_upper-.x$lnCVR_lower)/(2 * 1.96),
                                                       control = list(optimizer = "optim", 
                                                                      optmethod = "Nelder-Mead", 
                                                                      maxit = 1000),
                                                       verbose = F)))

plot3.female.meta.VR.all <- metacombo_female.plot3.VR.all %>%
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                          control = list(optimizer = "optim", 
                                         optmethod = "Nelder-Mead", 
                                         maxit = 1000), 
                          verbose = F)))

plot3.female.meta.RR.all <- metacombo_female.plot3.RR.all %>%
                            mutate(model_lnRR = map(data, 
                                                    ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                                                    control = list(optimizer = "optim", 
                                                                   optmethod = "Nelder-Mead", 
                                                                   maxit = 1000), 
                                                    verbose = F)))

plot3.female.meta.RR.all <- plot3.female.meta.RR.all %>% 
                            mutate(GroupingTerm = "All")

plot3.female.meta.CVR <- bind_rows(plot3.female.meta.CVR, plot3.female.meta.CVR.all)
 plot3.female.meta.VR <- bind_rows(plot3.female.meta.VR,   plot3.female.meta.VR.all)
 plot3.female.meta.RR <- bind_rows(plot3.female.meta.RR,   plot3.female.meta.RR.all)

plot3.female.meta.CVR.b <- as.data.frame(plot3.female.meta.CVR %>% 
                                           group_by(GroupingTerm) %>%
                                          mutate(    lnCVR = map_dbl(model_lnCVR, pluck(2)), 
                                               lnCVR_lower = map_dbl(model_lnCVR, pluck(6)),
                                               lnCVR_upper = map_dbl(model_lnCVR, pluck(7)), 
                                                  lnCVR_se = map_dbl(model_lnCVR, pluck(3)),
                                                  lnCVR_I2 = map_dbl(model_lnCVR, pluck("I2"))))[, c(1, 4:8)]

add.row.hearing <- as.data.frame(t(c("Hearing", NA, NA, NA, NA,NA))) %>% 
                                 setNames(names(plot3.female.meta.CVR.b))

plot3.female.meta.CVR.b <- bind_rows(plot3.female.meta.CVR.b, add.row.hearing)
plot3.female.meta.CVR.b <- plot3.female.meta.CVR.b[order(plot3.female.meta.CVR.b$GroupingTerm), ]

plot3.female.meta.VR.b <- as.data.frame(plot3.female.meta.VR %>% 
                                        group_by(GroupingTerm) %>%
                                        mutate(      lnVR = map_dbl(model_lnVR, pluck(2)), 
                                               lnVR_lower = map_dbl(model_lnVR, pluck(6)),
                                               lnVR_upper = map_dbl(model_lnVR, pluck(7)), 
                                                  lnVR_se = map_dbl(model_lnVR, pluck(3)),
                                                  lnVR_I2 = map_dbl(model_lnVR, pluck("I2"))))[, c(1, 4:8)]

>>>>

---
title: 'Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary
  and statistical implications'
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Alistair Senior,
  Daniel Noble, & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 6
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

```{r packages1, echo = FALSE, eval = TRUE, messages = FALSE}
library(pacman)
pacman::p_load(readr, dplyr,metafor, devtools, purrr, tidyverse, tidyr, tibble, kableExtra, robumeta, ggpubr, ggplot2, png, grid, here, knitr, pander, splus2R)
```

@@

This supplementary material follows the general structure described in our methods and is outlined here in Figure \@ref(fig:FigG1). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis. This is followed by how the first order and second order meta-analyses were conducted. Supplemental tables and figures that are referenced within the main manuscript can be found at the end of the document. Various parts of the workflow can be conveniently referenced using the table of contents panel (located along the side). We will refer back to this figure as we describe the various aspects of the code and data processing in relevant places.

```{r FigG1, fig.width = 4, fig.height = 8, fig.cap="Workflow of data process and meta-analysis"}

@@

calculate_population_stats <- function(mydata, min_individuals = 5){
    mydata %>% 
    group_by(population,
             production_center,  
             strain_name,  
             sex) %>%
    summarise(trait  = parameter_name[1],
              x_bar  = mean(data_point, na.rm = TRUE),
              x_sd   =   sd(data_point, na.rm = TRUE),
              n_ind  = n()) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%   
    group_by(population) %>%
    filter(all(c("male", "female") %in% sex)) %>% # Removes any entries where only 1 sex is present
    multi_spread(sex, c(x_bar, x_sd, n_ind))      # Spreads the data out into wide format, so sexes are beside each other which is more typical of meta-analytic data
}

@@

```
## Mean-variance relationships
We can also have a look at the mean-variance relationships in these data (Figure \@ref(fig:FigCor)). Note that these data have not been filtered to the same extent that the data is that is used in the meta-analyses, but this is simply here to demonstrate that mean-variance relationships exist and are strong providing strong justification for using the log coefficient of variation ratio (lnCVR).

```{r FigCor, fig.height=3, message = FALSE, fig.cap= "Mean-variance relationships (log(Mean) vs log(SD, standard deviation)) across all traits for males (A) and females (B)."}

# First, clean up the data

 results <- data %>%
            group_by(population,
             production_center,  
             strain_name,  
             parameter_name,
             sex) %>%
    summarise(trait  = parameter_name[1],
              x_bar  = mean(data_point, na.rm = TRUE),
              x_sd   =   sd(data_point, na.rm = TRUE),
              n_ind  = n()) %>%
    ungroup() %>%
    filter(n_ind > 5) %>%   # NOTE here that you are excluding data with 5 exactly....
    group_by(population) %>%
    filter(all(c("male", "female") %in% sex)) %>% # Removes any entries where only 1 sex is present
    multi_spread(sex, c(x_bar, x_sd, n_ind)) %>%
    mutate(lnMean_m = log(male_x_bar),
           logSD_m  = log(male_x_sd),
           lnMean_f = log(female_x_bar),
           lnSD_f   = log(female_x_sd)) %>%
    filter(!is.na(logSD_m) & !is.inf(logSD_m)) %>%
    filter(!is.na(lnSD_f) & !is.inf(lnSD_f))

male <- ggplot(results, aes(x = lnMean_m, y = logSD_m)) +
    geom_point(color = "blue", size = 2) +
   geom_smooth(method = "lm", se = FALSE, color="black") + 
    labs(x = "log (Mean)",
         y = "log (SD)") + 
    theme_classic()

female <- ggplot(results, aes(x = lnMean_f, y = lnSD_f)) +
    geom_point(color = "pink", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color="black")+
    labs(x = "log (Mean)",
         y = "log (SD)") + 
    theme_classic()

ggarrange(male, female, ncol = 2, nrow = 1, labels = c("A", "B"))

@@

We can see that there are strong relationships between the log transformed means and standard deviations for both males (Figure \@ref(fig:FigCor)A) and females (Figure \@ref(fig:FigCor)B), with the Person's correlation coefficient being `r round(stats::cor(results$lnMean_m,results$logSD_m, use = "complete.obs"), digits = 3) ` and `r round(cor(results$lnMean_f, results$lnSD_f, use = "complete.obs"), digits = 3) `, respectively.

@@

In this section, we cover the workflow described in Figure \@ref(fig:FigG1)C and D.

@@

This section covers Figure \@ref(fig:FigG1) C, by conducting a meta-analysis for all traits and producing the meta-analysed effect sizes (lnVR, lnCVR and lnRR).

@@

Here we cover Figure \@ref(fig:FigG1)D, removing non-coverging model results and non-informative traits. Procedure names, grouping variables and trait names ("parameter_names") are the merged back together with the results from the metafor analysis above.

@@

## Meta-analysis: condensing non-independent traits 
This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted meta analyses on these correlated parameters to collapse the number of levels. This section describes the workflow depicted in Figure \@ref(fig:FigG1)E & F.

### Collapsing and merging correlated parameters
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation. The workflow here relates to Figure \@ref(fig:FigG1)E.

@@

## Second-order meta-analysis for functional groups
Here, we describe the workflow depicted in Figure \@ref(fig:FigG1)H. We now use our uncorrelated traits from our first order meta-analysis to conduct a secondary meta-analysis by analysing the 9 functional trait groups. We estimate an overall mean lnCVR, lnVR and lnRR for each functional trait group.

@@

To see how much heterogeneity there is for each trait across different populations on average (for different functional groups and overall), we also conducted secondary meta-analyses by combining the total heterogeneities. More specifically, we combined the logarithm of the sum of the three variance components (Strain, Location and Unit; see above) with 1/2(N[effect size] - 1) as the sampling variance via the random effect model. That is, we conducted 3 sets of 10 secondary meta-analyses as with those of mean effect sizes for each trait, i.e. meta-analyzing log(total heterogeneities) for 9 functional groups and all the groups combined (Figure \@ref(fig:FigG1)I). The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses and specifically Figure \@ref(fig:FigG1)I & J. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

@@

Here we exclude traits without variation between mouse strains and merge data sets containing metafor results with procedure etc. names. Dealing with the correlated parameters, and running the second -order meta-analysis for heterogeneity data (Figure \@ref(fig:FigG1)I)

@@

# Meta-analysis results
## Preparing: Figure 3 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

## Main Manuscript: Figure 3

```{r Fig3, fig.cap = "Panel A shows the numbers of traits across functional groups that are either male-biased (blue-green) or female-biased (orange-red), as calculated in Figure 1.1D. Panel B shows effect sizes and 95% CI from separate meta-analysis for each functional group (step H in Figure1.1). Both panels represent results evaluated across all traits. Traits that are male biased are shown in blue, whereas female bias data is represented in orange."}
Fig4 <- ggarrange(malebias_Fig2_alltraits, Metameta_Fig3_alltraits,  nrow = 2, align = "v", heights = c(10, 9), labels = c("A", "B"))
Fig4
```
